
# Shaylee O {-}

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(here)
library(readr)
```


```{r}
trees <- read_csv(here("data", "trees1.csv"))

dim(trees)
head(trees)
tail(trees)

```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))


```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))
```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))|>
  head(5)
```

The top five most common tree SPECIES on campus are :
White spruce - 100, 
Crabapple - 78, 
Silver maple - 66, 
Hackberry - 65, 
Sugar maple - 64. 

then, top 10 are :
Honey locust - 55
Norway maple - 54
American elm - 31
Red maple - 31
Eastern redbud - 27 

Why is it that these trees are most common on campus? 100 White spruce is a lot of one tree. Does the sizing of the trees play any role in this? Because it is possible that if a tree species is very small than it could skew the number of them that are planted on campus because they take up lots of or very little room. 

```{r}
trees |>
  summarize(smallest = min(`Diameter at Breast Heigh (DBH)`), medium = mean(`Diameter at Breast Heigh (DBH)`), largest = max(`Diameter at Breast Heigh (DBH)`))
```
```{r}
trees |>
  select(`Diameter at Breast Heigh (DBH)`, SPECIES)|>
  arrange(desc(`Diameter at Breast Heigh (DBH)`))|>
  head(10)
```

```{r}
top_9_species <- trees |>
  filter(SPECIES %in% c("White spruce","Crabapple","Silver maple","Hackberry", "Sugar maple","Honey locust","Norway maple","American elm","Red maple"))|>
  select('Diameter at Breast Heigh (DBH)',SPECIES)
```

```{r}
species_counts <- top_9_species |>
  count(SPECIES, name = "Count")
```


```{r}
sum_top_9 <- top_9_species |>
  group_by(SPECIES) |>
  summarize(
    "smallest diameter" = min(`Diameter at Breast Heigh (DBH)`),
    "average diameter" = mean(`Diameter at Breast Heigh (DBH)`),
    "largest diameter" = max(`Diameter at Breast Heigh (DBH)`),
    .groups = "drop")
str(sum_top_9)
```

```{r}

sum_top_9_long <- sum_top_9 |>
  pivot_longer(cols = c(`smallest diameter`, `average diameter`, `largest diameter`),
               names_to = "Type",
               values_to = "Diameter")

```


```{r}
sum_top_9_long <- sum_top_9_long |>
  left_join(species_counts, by = "SPECIES") |>
  mutate(
    Type = factor(Type, levels = c("smallest diameter", "average diameter", "largest diameter")),
    SPECIES = fct_reorder(SPECIES, Count, .desc = TRUE))
```


```{r}
ggplot(sum_top_9_long,aes(x= SPECIES, y = Diameter, fill = Type))+
  geom_col(position = position_dodge())+
  coord_flip()+
  labs(title ="Campus Trees Diameter by Species", x = "Species (top 9)", y = "Diameter at Breast Height (inches)", fill = "Measurement Type", caption = "Data from Macalester Tree Inventory Drive (LAST UPDATED JULY 2024)")+
  theme_minimal()

```
```{r}
library(ggiraph)


```


```{r}
library(dplyr)
trees_clean <- trees %>% filter(!is.na(Memorial))

```

```{r}
trees_clean |>
  select(SPECIES, Memorial)

```


38 tree memorials!

```{r}
trees_clean |>
  group_by(SPECIES) |>
  summarise(Memorials = paste(Memorial, collapse = ", "), .groups = "drop")

```
Sugar maple - 5 memorials 
white oak - 3 memorials
Ohio buckeye - 3 
American Elm - 3
blue spruce - 3

```{r}
library(DataExplorer)

create_report(trees)

```

```{r}
introduce(trees)
```


