---
title: "EDA-Shaylee"
---

```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(here)
library(readr)
```


```{r}
trees <- read_csv(here("data", "trees1.csv"))

dim(trees)
head(trees)
tail(trees)

```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))


```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))
```

```{r}
trees |> 
  count(SPECIES)|>
  arrange(desc(n))|>
  head(5)
```

The top five most common tree SPECIES on campus are :
White spruce - 100, 
Crabapple - 78, 
Silver maple - 66, 
Hackberry - 65, 
Sugar maple - 64. 

then, top 10 are :
Honey locust - 55
Norway maple - 54
American elm - 31
Red maple - 31
Eastern redbud - 27 

Why is it that these trees are most common on campus? 100 White spruce is a lot of one tree. Does the sizing of the trees play any role in this? Because it is possible that if a tree species is very small than it could skew the number of them that are planted on campus because they take up lots of or very little room. 

```{r}
trees |>
  summarize(smallest = min(`Diameter at Breast Heigh (DBH)`), medium = mean(`Diameter at Breast Heigh (DBH)`), largest = max(`Diameter at Breast Heigh (DBH)`))
```
```{r}
trees |>
  select(`Diameter at Breast Heigh (DBH)`, SPECIES)|>
  arrange(desc(`Diameter at Breast Heigh (DBH)`))|>
  head(10)
```

```{r}
top_9_species <- trees |>
  filter(SPECIES %in% c("White spruce","Crabapple","Silver maple","Hackberry", "Sugar maple","Honey locust","Norway maple","American elm","Red maple"))|>
  select('Diameter at Breast Heigh (DBH)',SPECIES)
```


```{r}
sum_top_9 <- top_9_species |>
  group_by(SPECIES) |>
  summarize("smallest diameter" = min(`Diameter at Breast Heigh (DBH)`),"median diameter" = mean(`Diameter at Breast Heigh (DBH)`), "largest diameter" = max(`Diameter at Breast Heigh (DBH)`))
pivot_longer()
```

```{r}
ggplot(sum_top_9,aes(x= diameter, color= `smallest diameter`))+
  geom_point()+
  facet_wrap(SPECIES)+
  labs(title ="Species and Their Smallest, Average, and Largest Diameters on Campus", x = "Diameter", caption = "Made by Shaylee")
```


